---
<<<<<<< HEAD
title: "Reproducible Research Peer Assessment 1"
author: "Mihhail Fjodorov"
date: "Thursday, May 14, 2015"
output: html_document
---

### Loading and preprocessing the data

First load the data into R and convert date varible to date class
load dplyr, ggplot2 and stats packages
```{r echo = TRUE}
input <- read.csv("activity.csv", header = T, sep = ",")
input$date <- as.Date(input$date, "%Y-%m-%d")
require(dplyr,quietly = FALSE, warn.conflicts = FALSE)
require(ggplot2,quietly = FALSE, warn.conflicts = FALSE)
require(stats,quietly = FALSE, warn.conflicts = FALSE)
```

Calculate the number of steps per day, ignoring the missing values.
I'm using the dplyr package

```{r echo = TRUE}
stepPerDay <- input %>%
        group_by(date) %>%
        summarize(total_steps = sum(steps))%>%
        arrange(date)
```

### What is mean total number of steps taken per day?

Plot a histogram of total number of steps per day.

```{r echo = TRUE}
histStepDay <- ggplot(stepPerDay, aes(total_steps))
histStepDay <- histStepDay + geom_histogram(fill = "steelblue")
histStepDay <- histStepDay + ggtitle( "Number of Steps by Day" )
histStepDay <- histStepDay + xlab("Steps per Day")
histStepDay <- histStepDay + ylab("Number of Days")
histStepDay
```

The mean of the total number of steps taken per day
```{r echo = TRUE}
mean(stepPerDay$total_steps, na.rm = T)
```
The median of the total number of steps taken per day
```{r echo = TRUE}
median(stepPerDay$total_steps, na.rm = T)
```

### What is the average daily activity pattern?

The average number of steps taken over a 5 minute interval, averaged across all days.
```{r echo = TRUE}
fiveMinAve <- input %>%
        group_by(interval) %>%
        summarize(avg_step = mean(steps, na.rm = T)) %>%
        arrange(interval)
```

Create a Time Series plot of the Average number of Steps takes over a 5 minute interval.
```{r echo = TRUE}
ts <- ggplot(fiveMinAve, aes(x = interval, y = avg_step))
ts <- ts + geom_line(colour = "steelblue", size = 0.5)
ts <- ts + ggtitle("Average Steps over Interval")
ts + xlab("five minute interval") + ylab("Average number of Steps")
```

Which 5-minute interval, on average across all the days in the dataset contains the maximum number of steps?
```{r echo = TRUE}
as.data.frame(fiveMinAve[fiveMinAve$avg_step %in% max(fiveMinAve$avg_step),])
```

### Imputing missing values

Number of missing values in the data set (only steps varialbe considered)
```{r echo = TRUE}
sum(is.na(input$steps))
```
Filling in all of the missing values in the dataset. I chose to use the mean value for the
5 minute interval.


First join the input data with the 5-minute interval average
```{r echo = TRUE}
joinedData <- inner_join(input, fiveMinAve)
```
If the step value is NA replace it with the average value for the 5 minute interval
```{r echo = TRUE}
for(i in 1:dim(joinedData)[1]){
        if(is.na(joinedData[i,]$steps)){
                joinedData[i,]$steps <- joinedData[i,]$avg_step                
        }
}
```
Create a new dataset that is equal to the original dataset but with the missing data filled in
with the average values for the 5 minute intervals.
```{r echo = TRUE}
joinedData <-joinedData[,1:3]
```
Get the total number of steps per day
```{r echo = TRUE}
stepPerDay2 <- joinedData %>%
        group_by(date) %>%
        summarize(total_step = sum(steps)) %>%
        arrange(date)
```
Plot a histogram of total number of steps per day with missing values being replaced with averages.
```{r echo = TRUE}
histDay2 <- ggplot(stepPerDay2, aes(total_step))
histDay2 <- histDay2 + geom_histogram(fill = "steelblue")
histDay2 <- histDay2 + ggtitle( "Number of Steps by Day (NA's imputed)" )
histDay2 <- histDay2 + xlab("Steps per Day")
histDay2 <- histDay2 + ylab("Number of Days")
histDay2
```

The new mean of the total number of steps taken per day.
```{r echo = TRUE}
mean(stepPerDay2$total_step)
```
The new median of the total number of steps taken per day.
```{r echo = TRUE}
median(stepPerDay2$total_step)
```
There isn't a significat different between the mean/meadian claculated with NA's
and after the NA's were replaced with the average value of steps in the five minute interval.

### Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r echo = TRUE}
joinedData$weekday <- weekdays(joinedData$date, abbreviate = FALSE)
```
Create a factor variable which takes value 1 if the day is weekday and the value 2 if the day is weekend day.
```{r echo = TRUE}
joinedData$day_index <- 0
for(i in 1:dim(joinedData)[1]){
        if (joinedData[i,]$weekday %in% c(
                "Monday","Tuesday","Wednesday","Thursday","Friday" )){
                joinedData[i,]$day_index <- 1
        }else{
                joinedData[i,]$day_index <- 2    
        }
}
```
Coerce day_index into factor variable
```{r echo = TRUE}
joinedData$day_index <- as.factor(joinedData$day_index)
levels(joinedData$day_index) <- c("weekday", "weekend")
head(joinedData)
```
Calculate the average step per interval grouped by week/weekend day.
```{r echo = TRUE}
weekdayPlotData <- joinedData %>%
        select(steps, interval, day_index) %>%
        group_by(day_index, interval) %>%
        summarize(mean_step = mean(steps)) %>%
        arrange(day_index)

options(scipen = 999)
```
Create a panel Time Series plot which splits the mean number of steps over the five minute interval
according to weekdays and weekends.
```{r echo = TRUE}
panelPlot <- ggplot(weekdayPlotData, aes(x = interval, y = mean_step))
panelPlot <- panelPlot + geom_line(colour = "steelblue", size = 0.2)
panelPlot <- panelPlot + facet_wrap(~ day_index, nrow = 2, ncol = 1)
panelPlot <- panelPlot + ggtitle("Mean Number Steps per Interval (Weekdays/Weekends)")
panelPlot + xlab("Mean Number of Steps") + ylab("5 minute interval")
```
=======
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
